---
title: "The Game of floods"
author: "Andres Baeza"
date: "January 20, 2017"
output: html_document
---
This is a markdown document to describe how to generate scenarios for the game of flooding and scarcity In Mexico City. The scenarios correspond to the events of enviromental harmful events and the strategies of adaptation under uncertain climate signal and authority responses. Information about the game can be obtained in Shelton et al. (In preparation).

The game can be formalized mathematically. In this way we can aim to find the optimal strategy of decition that lead to the maximum outcome for a player, given the enviroment in which decision are made. We do so by creating scenarios of harful events over time. These scenarios are combined of number of events of different magnitude for two type of hazards: water scarcity and flooding.

###Load the R Packages
```{r}
require(ggplot2)#plots
require(gramEvol)#genetic algorithm
require(knitr) #tables
require(kableExtra) #tables
```

##Define the parameters of the game and the rules
###Define length of game
```{r}
years = 6 #number of years
ts_py = 4 #time-steps per year
length_total= years * ts_py  #length of the time series
```

###Define damage of environmental hazards and imcome per year 
In the game, three types of events can be defined based on the magnitude: moderate, grave and Critical such that:

$E=$ $\{e_1,e_t,e_3\}$
The level of damage caused by an event, $e$ is define as the cost of the event, such that 
$C=$ $\{C_{e_1},C_{e_2},C_{e_3}\}$

Every time-step, $t$ a player can take an action $a_j$ froma set of posible actions $A=$ $\{a_j\}$ for $j=$ $\{1,...,J\}$
The set of posible actions are





Each player recives a time-step income, $I_t$.


```{r}
cost_moderate_event=1#Moderate
cost_grave_event=4#Grave
cost_critical_event=5#Critical

Income=10

p_response=0.05 # likelihood authority responds to a protest
```

### Define the cost of each action and level of protection
```{r}
Cost_protest=3
Cost_WS1=1
Cost_WS2=5
Cost_HM1=1
Cost_HM2=4

protection_WS1=1
protection_WS2=2

protection_HM1=1
protection_HM2=2
```

###Table of action options, costs, and level of protection
```{r}

Table1<-data.frame(Actions=c("Do nothing","Protests","Storage water 1","Storage water 2","House modification 1","House modification 2"),
           Cost=c(0,Cost_protest,Cost_HM1,Cost_HM2,Cost_HM1,Cost_HM2),
           Protection=c(0,"Unknown","Moderate water shortage","Grave water shortage","Moderate flood","Grave flood"))

kable(Table1,"html") %>%
  column_spec(column = 2,width="5em") %>%
  column_spec(column = 3,width="10em")
```


###Define number of events with different damage per sequence of events
```{r}
#cost of water scarcity
number_critical_D=4
number_grave_D=2
number_moderate_D=1

#cost of floods
number_critical_W=4
number_grave_W=1
number_moderate_W=1


Total=number_critical_D+number_grave_D+number_moderate_D+number_critical_W+number_grave_W+number_moderate_W
if (Total > years * ts_py){
  stop() 
  print("stop: more events than timesteps")
}
```

#Define the data frame to store the information to define a scenario
Using the rules of the game and the parametrs of costs and protection
```{r, echo=TRUE,fig.height=3,fig.width=8}
scenarios<-data.frame(matrix(ncol=7,nrow=length_total))
colnames(scenarios)<-c("Environment","year","season","rounds","Event","cost_event","action-taked")
scenarios$season<-rep(c("D","D","W","W"),years)
scenarios$year<-rep(seq(1,years,1),each = ts_py)
scenarios$rounds<-rep(c(1,2,1,2), years)
scenarios$Event<-rep(0,length_total)
scenarios$cost_event<-rep(0,length_total)

#define events in dry season
scenarios$Environment<-rep("A",
                            length_total)


level_3_rounds<-sample(x = which(scenarios$season =="D"),
                       size = number_critical_D,
                       replace = F)

scenarios$Event[level_3_rounds]<-3
scenarios$cost_event[level_3_rounds]<-cost_critical_event


level_2_rounds<-sample(x = which(scenarios$season =="D" & scenarios$Event == 0),
                       size = number_grave_D,
                       replace = F)

scenarios$Event[level_2_rounds]<-2
scenarios$cost_event[level_2_rounds]<-cost_grave_event



level_1_rounds<-sample(x = which(scenarios$season =="D" & scenarios$Event ==0),
                       size = number_moderate_D,
                       replace = F)

scenarios$Event[level_1_rounds]<-1
scenarios$cost_event[level_1_rounds]<-cost_moderate_event

#define events in wet season
level_3_rounds<-sample(which(scenarios$season =="W"),size = number_critical_W,replace = F)
scenarios$Event[level_3_rounds]<-3
scenarios$cost_event[level_3_rounds]<-cost_critical_event

level_2_rounds<-sample(which(scenarios$season =="W" & scenarios$Event == 0),size = number_grave_W,replace = F)
scenarios$Event[level_2_rounds]<-2
scenarios$cost_event[level_2_rounds]<-cost_grave_event

level_1_rounds<-sample(which(scenarios$season =="W" & scenarios$Event ==0),size = number_moderate_W,replace = F)
scenarios$Event[level_1_rounds]<-1
scenarios$cost_event[level_1_rounds]<-cost_moderate_event

```

#Optimization using GA
A scheme to evaluate what sequence of actions (the strategy) is the one that generates the largest benefit is constructed using Genetic algorithm (GA).
First, an objective function is contructed to compute the time-varying outcome of income, costas and protection using:

The goal is to find the sequence of actions, $S=$$\{a_1,...,a_t,...,a_T\}$, that maximized outcome $\sum_{t=1}^{T}O_t$. Formally, the timestep outcome is calculated using 

$$O_t = I_t - (C_{e,t} - P_{a,t})$$  

where
$I_t$: Income per unit of time

$C_{e,t}$: Cost of an event of magnitude $e$ at time $t$ (if any)

$P_{a,t}$: The level of protection of action $a$ at time $t$ (if any) 

\begin{equation*}
\begin{aligned}
& \underset{S}{\text{maximize}}
& & \sum_{t=1}^{T}O_t \\
& \text{subject to}
& & S \leq b_i, \; i = 1, \ldots, m.
\end{aligned}
\end{equation*}



```{r, echo=TRUE}

evalFunc <- function(ch) {

  O<-numeric(1+length(ch)) #length of gen based on number of "codons"

    
  O[1]=0  #nothing in the first time-step
  cost_Ad=0
  
   for(t in 2 :(1 + length(ch))){
     cost_E=scenarios$cost_event[(t-1)] 
 
      if (t <= 4){
       if (scenarios$season[(t-1)] == "D"){
        if (scenarios$Event[(t-1)] == 1){if (any(ch[1:(t-1)] == 3 | ch[1:(t-1)] == 4)){cost_E=0}}
        if (scenarios$Event[(t-1)] == 2){if (any(ch[1:(t-1)]== 4)){cost_E=0}}
       }

       
       if (scenarios$season[(t-1)] == "W"){
        if (scenarios$Event[(t-1)] == 1){if (any(ch[1:(t-1)] == 5 | ch[1:(t-1)] == 6)){cost_E=0}}
        if (scenarios$Event[(t-1)] == 2){if (any(ch[1:(t-1)] == 6)){cost_E=0}}
       }
      }
 #########################################################################################    
     
     if (t > 4){
        if (scenarios$season[(t-1)] == "D"){
          if (scenarios$Event[(t-1)] == 1){if (any(ch[(t-4):(t-1)] == 3) | any(ch[(t-4):(t-1)] == 4)){cost_E=0}}
          if (scenarios$Event[(t-1)] == 2){if (any(ch[(t-4):(t-1)] == 4)){cost_E=0}}
        }

        if (scenarios$season[(t-1)] == "W"){
          if (scenarios$Event[(t-1)] == 1){if (any(ch[(t-4):(t-1)] == 5) | any(ch[(t-4):(t-1)] == 6)){cost_E=0}}
          if (scenarios$Event[(t-1)] == 2){if (any(ch[(t-4):(t-1)] == 6)){cost_E=0}}
        }  
      }
 #########################################################################################    
# considert the cost of critical events     
      if (scenarios$Event[(t-1)] == 3){cost_E=3} 
 #########################################################################################    
 # Calculate the cost of the actions      
     if (ch[(t-1)]==1){cost_Ad=0}
     if (ch[(t-1)]== 2){cost_Ad=Cost_protest * rbinom(n = 1,size=1,prob = (1-p_response))}
     if (ch[(t-1)]==3){cost_Ad=Cost_WS1}
     if (ch[(t-1)]==4){cost_Ad=Cost_WS2}
     if (ch[(t-1)]==5){cost_Ad=Cost_HM1}
     if (ch[(t-1)]==5){cost_Ad=Cost_HM2}

    O[t] = Income - cost_E - cost_Ad  #objective function
     
}


 current_solution_survivalpoints= (- sum(O[-1]))

 return(current_solution_survivalpoints)
}

#Monitor function
monitorFunc <- function(result) {
    cat("Best of gen: ", min(result$best$cost), "\n")
}
```


```{r, echo=TRUE}

best_strategy<-EvolutionStrategy.int(genomeLen = length_total, 
                                      genomeMin = rep(1,length_total), 
                                      genomeMax = rep(6,length_total), 
                                      evalFunc = evalFunc,iterations = 400)


best_strategies<-data.frame(genome=best_strategy$best$genome,
                            simul=rep(1,length_total),
                            timestep=1:length_total)
for (i in 2:10){
    best_strategy<-EvolutionStrategy.int(genomeLen = length_total, 
                                      genomeMin = rep(1,length_total), 
                                      genomeMax = rep(6,length_total),
                                      evalFunc = evalFunc,
                                      iterations = 400)

    best_strategies=rbind(best_strategies,cbind(genome=best_strategy$best$genome,simul=rep(i,length_total),timestep=1:length_total))
print(evalFunc(subset(best_strategies,simul==i)$genome))
    }
```


##Display the environmental Scenario
red and blue dots for Dry and Wet seasons respectivelly 
```{r, echo=TRUE}
ggplot(data = scenarios,aes(y=Event,x=(1:length_total),size=Event,colour=season))+geom_point()
```

##Display the frequqncy of Evolving actions
Counts of actions used  in the environmental scenario
```{r,fig.align="center",fig.width=6,fig.height=3}
ggplot(data=best_strategies,aes(x=factor(genome),colour=factor(genome)))+geom_bar()

```


```{r,fig.align="center",fig.width=6,fig.height=3}
ggplot(data=best_strategies,aes(y=timestep,x=factor(genome),colour=factor(genome)))+geom_boxplot()
```
