---
title: "The Game scenariosC"
author: "Andres Baeza"
date: "January 20, 2017"
output: html_document
---
This is a markdown document to describe how to generate scenarios hydrological harmful events of flooding and scarcity
```{r}
require(ggplot2)
```
#Scenario A
```{r, echo=TRUE,fig.height=3,fig.width=8}
length_total=30
scenariosA<-data.frame(matrix(ncol=7,nrow=length_total))
colnames(scenariosA)<-c("Environment","year","season","rounds","Event","cost_event","action-taked")
scenariosA$season<-rep(c("D","D","D","W","W","W"),5)
scenariosA$year<-rep(seq(1,5,1),each =6)
scenariosA$rounds<-rep(c(1,2,3),10)
scenariosA$Event<-rep(0,length_total)
scenariosA$cost_event<-rep(0,length_total)

#scenario A
#define events in dry season
scenariosA$Environment<-rep("A",length_total)
level_3_rounds<-sample(which(scenariosA$season =="D"),size = 5,replace = F)
scenariosA$Event[level_3_rounds]<-3
scenariosA$cost_event[level_3_rounds]<-3
level_2_rounds<-sample(which(scenariosA$season =="D" & scenariosA$Event == 0),size = 3,replace = F)
scenariosA$Event[level_2_rounds]<-2
scenariosA$cost_event[level_2_rounds]<-2
level_1_rounds<-sample(which(scenariosA$season =="D" & scenariosA$Event ==0),size = 2,replace = F)
scenariosA$Event[level_1_rounds]<-1
scenariosA$cost_event[level_1_rounds]<-1

#define events in wet season
level_3_rounds<-sample(which(scenariosA$season =="W"),size = 1,replace = F)
scenariosA$Event[level_3_rounds]<-3
scenariosA$cost_event[level_3_rounds]<-3

level_2_rounds<-sample(which(scenariosA$season =="W" & scenariosA$Event == 0),size = 2,replace = F)
scenariosA$Event[level_2_rounds]<-2
scenariosA$cost_event[level_2_rounds]<-2

level_1_rounds<-sample(which(scenariosA$season =="W" & scenariosA$Event ==0),size = 7,replace = F)
scenariosA$Event[level_1_rounds]<-1
scenariosA$cost_event[level_1_rounds]<-1

```

#Scenario B
```{r, echo=TRUE}

scenariosB<-data.frame(matrix(ncol=7,nrow=length_total))
colnames(scenariosB)<-c("Environment","year","season","rounds","Event","cost_event","action-taked")
scenariosB$season<-rep(c("D","D","D","W","W","W"),5)
scenariosB$year<-rep(seq(1,5,1),each =6)
scenariosB$rounds<-rep(c(1,2,3),10)
scenariosB$Event<-rep(0,length_total)
scenariosB$cost_event<-rep(0,length_total)

scenariosB$Environment<-rep("B",length_total)
#define events in dry season
level_3_rounds<-sample(which(scenariosB$season =="D"),size = 3,replace = F)
scenariosB$Event[level_3_rounds]<-3
scenariosB$cost_event[level_3_rounds]<-3
level_2_rounds<-sample(which(scenariosB$season =="D" & scenariosB$Event == 0),size = 3,replace = F)
scenariosB$Event[level_2_rounds]<-2
scenariosB$cost_event[level_2_rounds]<-2
level_1_rounds<-sample(which(scenariosB$season =="D" & scenariosB$Event ==0),size = 4,replace = F)
scenariosB$Event[level_1_rounds]<-1
scenariosB$cost_event[level_1_rounds]<-1
#define events in wet season
level_3_rounds<-sample(which(scenariosB$season =="W"),size = 3,replace = F)
scenariosB$Event[level_3_rounds]<-3
scenariosB$cost_event[level_3_rounds]<-3
level_2_rounds<-sample(which(scenariosB$season =="W" & scenariosB$Event == 0),size = 3,replace = F)
scenariosB$Event[level_2_rounds]<-2
scenariosB$cost_event[level_2_rounds]<-2
level_1_rounds<-sample(which(scenariosB$season =="W" & scenariosB$Event ==0),size = 3,replace = F)
scenariosB$Event[level_1_rounds]<-1
scenariosB$cost_event[level_1_rounds]<-1

```



#Scenario C
```{r, echo=TRUE}
scenariosC<-data.frame(matrix(ncol=7,nrow=length_total))
colnames(scenariosC)<-c("Environment","year","season","rounds","Event","cost_event","action-taked")
scenariosC$season<-rep(c("D","D","D","W","W","W"),5)
scenariosC$year<-rep(seq(1,5,1),each =6)
scenariosC$rounds<-rep(c(1,2,3),10)
scenariosC$Event<-rep(0,length_total)
scenariosC$cost_event<-rep(0,length_total)

scenariosC$Environment<-rep("C",length_total)
#define events in dry season
level_3_rounds<-sample(which(scenariosC$season =="D"),size = 1,replace = F)
scenariosC$Event[level_3_rounds]<-3
scenariosC$cost_event[level_3_rounds]<-3
level_2_rounds<-sample(which(scenariosC$season =="D" & scenariosC$Event == 0),size = 2,replace = F)
scenariosC$Event[level_2_rounds]<-2
scenariosC$cost_event[level_2_rounds]<-2
level_1_rounds<-sample(which(scenariosC$season =="D" & scenariosC$Event ==0),size = 7,replace = F)
scenariosC$Event[level_1_rounds]<-1
scenariosC$cost_event[level_1_rounds]<-1
#define events in wet season
level_3_rounds<-sample(which(scenariosC$season =="W"),size = 5,replace = F)
scenariosC$Event[level_3_rounds]<-3
scenariosC$cost_event[level_3_rounds]<-3
level_2_rounds<-sample(which(scenariosC$season =="W" & scenariosC$Event == 0),size = 3,replace = F)
scenariosC$Event[level_2_rounds]<-2
scenariosC$cost_event[level_2_rounds]<-2
level_1_rounds<-sample(which(scenariosC$season =="W" & scenariosC$Event ==0),size = 2,replace = F)
scenariosC$Event[level_1_rounds]<-1
scenariosC$cost_event[level_1_rounds]<-1


```



```{r, echo=TRUE}

table_concatScenarios<-cbind(scenariosA$season,scenariosA$Event,scenariosB$season,scenariosB$Event,scenariosC$season,scenariosC$Event)
colnames(table_concatScenarios)<-c("SA","EA","SB","EB","SC","EC")
write.csv(x = table_concatScenarios,file = "~/MEGADAPT/games/table_scenarios.csv")
```



```{r, echo=TRUE}
ggplot(data = scenariosA,aes(y=Event,x=(1:30)))+geom_line()+
  geom_line(data = scenariosB,colour="green")+
geom_line(data = scenariosC,colour="blue")
```

#Best strategy
A scheme to evaluate what sequence of actions, or the strategy that generate the largest benefits, measured as
$B = I - (C - A_d)-C_{A_{d}}$  
where
$C$: is the cost of the harful event
$A_d$: is the level of adaptation 

The goal is to find strategy $S$ that maximized benefit $B$

codon1=Do nothing
codon2=protests
codon3=water_Storage1
codon4=water_Storage2
codon5=house modification 1
codon6=house modification 2




```{r, echo=TRUE}

evalFunc <- function(ch) {

  B<-numeric(1+length(ch))

    
  B[1]=3
  cost_Ad=0
  
   for(t in 2 :(1 + length(ch))){
     cost_E=scenariosC$cost_event[(t-1)] 
 #########################################################################################    
     if (t <= 4){
      if (scenariosC$season[(t-1)] == "D"){
        if (scenariosC$Event[(t-1)] == 1){if (any(ch[1:(t-1)] == 3)){cost_E=0}}
        if (scenariosC$Event[(t-1)] == 2){if (any(ch[1:(t-1)] == 3) & any(ch[1:(t-1)] == 4)){cost_E=0}}
      }

      if (scenariosC$season[(t-1)] == "W"){
        if (scenariosC$Event[(t-1)] == 1){if (any(ch[1:(t-1)] == 5)){cost_E=0}}
        if (scenariosC$Event[(t-1)] == 2){if (any(ch[1:(t-1)] == 5) & any(ch[1:(t-1)] == 6)){cost_E=0}
          
          }
      
      }
       if(any(ch[1:(t-1)] ==3) & any(ch[1:(t-1)] ==4)){
        if(which(ch[1:(t-1)] == 4)[1] < which(ch[1:(t-1)] == 3)[1]){cost_E=Inf}
       }
       if(any(ch[1:(t-1)] ==5) & any(ch[1:(t-1)]==6)){
         if(which(ch[1:(t-1)] == 6)[1] < which(ch[1:(t-1)] == 5)[1]){cost_E=Inf}
      }
     }
 #########################################################################################    
     
     if (t > 4){
      if (scenariosC$season[(t-1)] == "D"){
        if (scenariosC$Event[(t-1)] == 1){if (any(ch[(t-4):(t-1)] == 3)){cost_E=0}}
        if (scenariosC$Event[(t-1)] == 2){if (any(ch[(t-4):(t-1)] == 3) & any(ch[(t-4):(t-1)] == 4)){cost_E=0}}
      }

      if (scenariosC$season[(t-1)] == "W"){
        if (scenariosC$Event[(t-1)] == 1){if (any(ch[(t-4):(t-1)] == 5)){cost_E=0}}
        if (scenariosC$Event[(t-1)] == 2){if (any(ch[(t-4):(t-1)] == 5) & any(ch[(t-4):(t-1)] == 6)){cost_E=0}}
      }  
       
       
       if(any(ch[(t-4):(t-1)] ==3) & any(ch[(t-4):(t-1)] ==4)){
        if(which(ch[(t-4):(t-1)] == 4)[1] < which(ch[(t-4):(t-1)] == 3)[1]){cost_E=Inf}
       }
       if(any(ch[(t-4):(t-1)] ==5) & any(ch[(t-4):(t-1)]==6)){
         if(which(ch[(t-4):(t-1)] == 6)[1] < which(ch[(t-4):(t-1)] == 5)[1]){cost_E=Inf}
   }
       
     }
 #########################################################################################    
     
     
     if (scenariosC$Event[(t-1)] == 3){
       cost_E=3
     } 
     
     
     if (ch[(t-1)]== 1){cost_Ad=0}
     if (ch[(t-1)]== 2){cost_Ad=1}
     if (ch[(t-1)] != 1 & ch[(t-1)]!= 2){cost_Ad=3}
      
    B[t] = B[t-1] + 2 - cost_E - cost_Ad
     
}


 current_solution_survivalpoints= - sum(B[-1])

 return(current_solution_survivalpoints)
}

#Monitor function
monitorFunc <- function(result) {
    cat("Best of gen: ", min(result$best$cost), "\n")
}
chrom<-c(3, 2, 1, 5, 3, 4, 2, 2, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 5, 6, 1, 2, 2, 3, 1, 4, 1, 2, 1)
evalFunc(chrom)
```


```{r, echo=TRUE}
require(gramEvol)
best_strategyC<-EvolutionStrategy.int(genomeLen = length_total, genomeMin = rep(1,length_total), genomeMax = rep(6,length_total), evalFunc = evalFunc,iterations = 400)
best_strategies<-data.frame(genome=best_strategyC$best$genome,simul=rep(1,30),timestep=1:30)
for (i in 2:10){
best_strategyC<-EvolutionStrategy.int(genomeLen = length_total, genomeMin = rep(1,length_total), genomeMax = rep(6,length_total), evalFunc = evalFunc,iterations = 400)
best_strategies=rbind(best_strategies,cbind(genome=best_strategyC$best$genome,simul=rep(i,30),timestep=1:30))
}
ggplot(data=best_strategies,aes(x=timestep,y=genome,colour=factor(simul)))+geom_point()+geom_smooth(method = "lm")
ggplot(data=best_strategies,aes(x=factor(genome),colour=factor(genome)))+geom_boxplot()
```

